Manuscript 1 code

### 2025-03-17 manuscript_1_code.Rmd
implemented plot aesthetic changes

## Figure formatting
```{r figure formatting info}
trends_theme <- theme_bw() + 
  theme(axis.title.x = element_text(size=24), 
        axis.text.x = element_text(size = 22, angle = 45, hjust = 1),
        axis.title.y = element_text(size=24), 
        axis.text.y = element_text(size = 22),
        plot.title = element_text(size = 24),
        panel.background = element_rect(fill = "white"), 
        panel.grid.major = element_blank(),  #remove major-grid labels
        panel.grid.minor = element_blank(),  #remove minor-grid labels
        plot.background = element_rect(fill = "white"),
        legend.text = element_text(size = rel(1)), # increase legend text
        legend.title = element_blank(),
        legend.position = 'bottom', # move legend to bottom
        legend.box = 'horizontal', # make legend horizontal
        legend.box.background = element_rect(colour = "black"))

model_themes <-   theme_bw() + 
  theme(axis.title.x = element_text(size=24), 
        axis.title.y = element_text(size=24), 
        axis.text.x = element_text(size=22), 
        axis.text.y = element_text(size=22),
        legend.text = element_text(size=24),
        plot.title = element_text(size=26), 
        panel.background = element_blank(), 
        panel.grid.major = element_blank(),  #remove major-grid labels
        panel.grid.minor = element_blank(),  #remove minor-grid labels
        plot.background = element_rect(fill = "white"),
        legend.title = element_blank(),
        legend.position = 'bottom', # move legend to bottom
        legend.box = 'horizontal', # make legend horizontal
        legend.box.background = element_rect(colour = "black"))  # add box around legend

@@

```{r Figure 1, fig.height = 10, fig.width=14}
plot_list <- lapply(levels(figure_1_data$habitat), function(hab) {
  ggplot(data = subset(figure_1_data, habitat == hab)) +
    geom_point(aes(x = year, y = percent_cover, colour = functional_group), size = 3) +
    geom_errorbar(aes(x = year, ymin = percent_cover - se, ymax = percent_cover + se, colour = functional_group), 
                  width = 0, size = 1.3) +
    geom_line(aes(x = year, y = percent_cover, colour = functional_group), linewidth = 1) +
    annotate("segment", x = 2006, xend = 2010, y = 75, yend = 75, linewidth = 2, colour = "grey") +
    geom_vline(xintercept = 2010, linetype = 2, colour = "black", size = 1) +
    geom_vline(xintercept = 2019, linetype = 2, colour = "red", size = 1) +
    scale_colour_manual(values = functional_group_colours) +
    scale_x_continuous(breaks = unique(figure_1_data$year)) +
    labs(y = "Percent Cover",
         x = "Year", title = "") +  
    trends_theme
})

@@

```{r figure 3 site supplement, fig.height = 8, fig.width = 16}
(figure_3_supplement <- ggarrange(figure_3d, figure_3e,  
                      ncol = 2, nrow = 1, 
                      common.legend = TRUE,
                      labels = c("a.", "b."),
                      legend = "bottom", 
                      font.label = list(size = 26, color = "black", face = "plain")))

>>>>

## Figure formatting
```{r figure formatting info}
trends_theme <- theme_bw() + 
  theme(axis.title.x = element_text(size=24), 
        axis.text.x = element_text(size = 22, angle = 45, hjust = 1),
        axis.title.y = element_text(size=24), 
        axis.text.y = element_text(size = 22),
        plot.title = element_text(size = 24),
        panel.background = element_rect(fill = "white"), 
        panel.grid.major = element_blank(),  #remove major-grid labels
        panel.grid.minor = element_blank(),  #remove minor-grid labels
        plot.background = element_rect(fill = "white"),
        legend.text = element_text(size=24),
        legend.title = element_blank(),
        legend.position = 'bottom', # move legend to bottom,
        legend.key.size = unit(2, 'cm'), # increase legend size
        legend.box = 'horizontal', # make legend horizontal
        legend.box.background = element_rect(colour = "black"))

model_themes <-   theme_bw() + 
  theme(axis.title.x = element_text(size=24), 
        axis.title.y = element_text(size=24), 
        axis.text.x = element_text(size=22), 
        axis.text.y = element_text(size=22),
        legend.text = element_text(size=24),
        plot.title = element_text(size=26), 
        panel.background = element_blank(), 
        panel.grid.major = element_blank(),  #remove major-grid labels
        panel.grid.minor = element_blank(),  #remove minor-grid labels
        plot.background = element_rect(fill = "white"),
        legend.title = element_blank(),
        legend.key.size = unit(2, 'cm'), # increase legend size
        legend.position = 'bottom', # move legend to bottom
        legend.box = 'horizontal', # make legend horizontal
        legend.box.background = element_rect(colour = "black"))  # add box around legend

@@

```{r Figure 1, fig.height = 10, fig.width=14}
plot_list <- lapply(levels(figure_1_data$habitat), function(hab) {
  ggplot(data = subset(figure_1_data, habitat == hab)) +
    geom_point(aes(x = year, y = percent_cover, colour = functional_group), size = 3) +
    geom_errorbar(aes(x = year, ymin = percent_cover - se, ymax = percent_cover + se, colour = functional_group), 
                  width = 0, size = 1.3) +
    geom_line(aes(x = year, y = percent_cover, colour = functional_group), linewidth = 1) +
    annotate("segment", x = 2006, xend = 2010, y = 75, yend = 75, linewidth = 2, colour = "grey") +
    geom_vline(xintercept = 2010, linetype = 2, colour = "black", size = 1) +
    geom_vline(xintercept = 2019, linetype = 2, colour = "red", size = 1) +
    scale_colour_manual(values = functional_group_colours) +
    scale_x_continuous(breaks = seq(2007, 2023, 2)) +
    labs(y = "Percent Cover",
         x = "Year", title = "") +  
    trends_theme
})

@@

```{r figure 3 site supplement, fig.height = 8, fig.width = 16}
(figure_3_supplement <- ggarrange(figure_3d, figure_3e,  
                      ncol = 2, nrow = 1, 
                      common.legend = TRUE,
                      labels = c("a.", "b."),
                      legend = "bottom", 
                      legend.grob = get_legend(figure_3a),
                      font.label = list(size = 26, color = "black", face = "plain")))
